---
title: "People's Happiness Based on Covid Cases & Climate Awareness"
author: "Group 3"
date: "November 12th, 2024"
output:
  html_document:
    df_print: paged
---

Report contributors:

-   Noya Barak (169097527)

-   Sara Haifa (169087012)

-   Satinder Kaur (169109308)

-   Spencer Mozeg (169099531)

-   Khushi Satra, (203383100)

```{r setup, include=FALSE}
# echo = FALSE will set the Rmd to *not* show the R code. Don't change this.
# You may change the default figure width and figure height as you please.
knitr::opts_chunk$set(echo = FALSE, message = FALSE, fig.width = 6)

# Put any libraries that you need to load here.
# DO NOT PUT "install.packages()" IN AN RMD FILE!!!
library(tidyverse)
library(arrow)
library(knitr)
```

# Instructions

You will only submit the PDF version of this document. To knit to PDF, you'll need to run `install.packages("tinytex")` in the console, followed by `tinytex::install_tinytex()` (DO NOT PUT THESE COMMANDS IN AN RMD FILE!!!). If you encounter errors in "Knit to PDF", you can "knit to html" and then print the html file to PDF using your operating system's PDF view (e.g. Adobe Acrobat). Only standalone PDF files will be accepted by MyLS.

# Abstract

The analysis presented is a continuation of our exploration of Disparate Data: Part 1. We explore how three distinct data sets(aware covid data, happiness data and climate awareness data) to explore different connections between each data set like examining the well-being, environmental awareness covid awareness. Firstly, the happiness data set shows the social-economic and well-being across different countries. The aware covid data talks about the awareness of covid. And lastly, the climate awareness data consists of the awareness in people across the world regarding the climate changes. The purpose of this analysis is to discover the relationship between happiness, awareness of climate change and covid data.

# Introduction

In 2020, global events such as the COVID-19 pandemic reshaped societal priorities, affecting public health, economic stability, and general well-being worldwide. This study explores the interrelations between national happiness levels, public awareness of climate change, and COVID-19 case data to understand broader implications on public sentiment and awareness, crossing over from the Covid 2020 data set, and Climate Awareness, and Happiness.

We imported and tidied the data sets using techniques to ensure they combined seamlessly. The `dplyr` package was used for data manipulation (e.g., `filter()`, `mutate()`, `group_by()`), and `pivot_longer()`/`pivot_wider()` for reshaping. We standardized column names with `janitor::clean_names()` and handled missing values with `na.rm = TRUE`. Finally, the cleaned data was exported using `write_parquet().`

In this report, we are going to explore some aspects climate change and the impact and/or perceptions of it by using exploratory techniques.

By the end of this report, we will show how climate awareness, economic stability, and low corruption support well-being and resilience.

# Data Description

## 1. Climate Awareness

```{r load_data1}
climate_opinion_address <- "https://data.humdata.org/dataset/dc9f2ca4-8b62-4747-89b1-db426ce617a0/resource/6041db5f-8190-47ff-a10b-9841325de841/download/climate_change_opinion_survey_2022_aggregated.xlsx"
climate_sheet_names <- climate_opinion_address |>
    loadWorkbook() |>
    names()
aware_sheet_name <- "climate_awareness"
climate_awareness_update2 <- climate_opinion_address |>
    read.xlsx(
        sheet = aware_sheet_name ) |>
    pivot_longer(
        cols = !contains(aware_sheet_name),
        names_to = "country",
        values_to = "score") |>
    mutate(
        climate_awareness = case_when(
            climate_awareness == "I have never heard of it" ~ "aware_no",
            climate_awareness == "I know a little about it" ~ "aware_alittle",
            climate_awareness == "I know a moderate amount about it" ~
                "aware_moderate",
            climate_awareness == "I know a lot about it" ~ "aware_alot",
            climate_awareness == "Refused" ~ "aware_refuse",
            climate_awareness == "(Unweighted Base)" ~ "aware_base" )) |>
    rename(answer = climate_awareness) 
climate_awareness_update2
climate_awareness<- climate_awareness_update2|>
    pivot_wider(
        names_from = answer,
        values_from = score)
write_parquet(climate_awareness, "climate_awareness.parquet")
climate_awareness
```

The data come from [a climate awareness surve](https://data.humdata.org/dataset/dc9f2ca4-8b62-4747-89b1-db426ce617a0/resource/6041db5f-8190-47ff-a10b-9841325de841/download/climate_change_opinion_survey_2022_aggregated.xlsx)y [^1]and describes countries awareness of climate change, grouping them into people who are knowledgeable about climate change to people who's never heard of it. It comes from an Excel document with other related climate topics, brought from the [2020 survey](https://dataforgood.facebook.com/dfg/tools/climate-change-opinion-survey) [^2]. After loading in the data from the Excel sheet, we separated the country and score column and simplified the survey answers to clearer headings. After loading in the data from the Excel sheet, pivoting was used to reshape the data from a wide format (with each survey response category in separate columns) to a long format that is easier to analyze the score of countries. Then, we renamed the survey options to shorter codes simplifies the data for future processing and analysis. Using "aware_no" instead of lengthy descriptions such as "I have never heard of it" makes the code more concise and readable. The climate_awareness column was renamed as well to ensure clarity and consistency in column naming. Next, we restructured the data so that survey response categories are column names (e.g., aware_no, aware_alittle) and their respective scores are values, making it easier to analyze each country's awareness level. Finally, the cleaned data was saved as a a.parquet file.

[^1]: <https://data.humdata.org/dataset/dc9f2ca4-8b62-4747-89b1-db426ce617a0/resource/6041db5f-8190-47ff-a10b-9841325de841/download/climate_change_opinion_survey_2022_aggregated.xlsx>

[^2]: <https://dataforgood.facebook.com/dfg/tools/climate-change-opinion-survey>

## 2. COVID-19 2020

```{r load_data2}
owid_address <- "https://covid.ourworldindata.org/data/owid-covid-data.csv"
covid_cases <- owid_address |>
    read_csv(
        col_types = cols(
            .default = col_double(),
            date = col_date(format = ""),
            iso_code = col_character(),
            location = col_character(),
            continent = col_character(),
            tests_units = col_character() ),
        progress = FALSE,
        show_col_types = FALSE )
covid_cases
covid_2020 <- covid_cases|>
  filter(year(date) == 2020 )|>
  group_by(country = location, continent) |>
  summarise(total_cases = sum(total_cases, na.rm = TRUE),
            .groups = "drop")|>
  select(country, continent, total_cases)
covid_2020
write_parquet(covid_2020, "covid_2020.parquet")
```

Collected from [here](https://covid.ourworldindata.org/data/owid-covid-data.csv)[^3], this data set showcases the cases in each country (as well as its continent), reported in 2020. It contains a wide range of information as well (like life expectancy, total_deaths, etc.)

[^3]: <https://covid.ourworldindata.org/data/owid-covid-data.csv>

When we tidied up the data, we focused specifically on Covid cases. Firstly, after loading out the data, we defined the columns with various functions, to ensure correct data types for example. Then we filtered by year as every collection was from the same year but different periods of 2020, which just made the data more cleaner. Next, we grouped location and continent to connect the pairs, summarized total cases to sum up the cases, and removed NA to ensure correct values. After that, we selected country, continent, and total_cases as those were our main columns we wanted to focus on.

## 3. World Happiness Report Score (from 2023 Report)

```{r load_data3}
happiness <- read.xlsx("DataForTable2.1.xlsx") |>
    janitor::clean_names() |>
    rename(
      country = 'country_name' ) |>
  filter(
    !is.na(life_ladder)) |>
  group_by(country) |>
  slice_max(year, n=1)

happiness
```

This dataset was taken from [Home \| The World Happiness Report](https://worldhappiness.report/) [^4] (specifically used from [here](https://worldhappiness.report/ed/2023/#appendices-and-data) [^5]) and records the areas of 'happiness' in each country. The column closest to this measure is the `life_ladder` column; happiness in each country.

[^4]: <https://worldhappiness.report/>

[^5]: <https://worldhappiness.report/ed/2023/#appendices-and-data>

In Part 1, cleaning up life_ladder was focused on. Tidying up this column had the following steps. `Janitor` used to clean up the names (turns uppercase into lower case and replacing spaces or such with underscores), renaming the country column (from `country_name`), filtering out the missing/incomplete values (`NA's`), and lastly only focusing on recent observations.

# Combining the Data

**Climate Awareness and Happiness data:**

1.  We used the inner_join function by country.
2.  Then we renamed the answer column to awareness_level
3.  We arrange the data in descending order by log_gdp_per_capita and then we got the top 5 countries with highest gdp.

**Happiness and Covid 2020 data:**

1.  We performed the combination of the data through the inner_join function
2.  This function helps recognize which countries are mutual between the two datasets, and therefore we do not have any missing data per country due to merging
3.  The only column that we modified was the year column, since it only provided covid statistics per individual year, rather than reporting a statistic of the net cases per country
4.  The plot that we made thus combined information from both the covid dataset (continent data and total covid cases) and data from the happiness dataset (gdp per capita and corruption rates)
5.  Through this, we can see how countries who overall generate less capital per person's population is affected by covid, and also if certain continents in the world's covid rates were higher

**Aware_Covid_Happiness**

1.  The aware_covid dataset was cleaned and combined in part 1 using multiple datasets but mainly covid_2020 and climate_awaress. It was combined by the country names present in both the date sets.
2.  This combined data set aware_covid consisting of column covid cases per capita calulated by dividing total cases per country by it's total population, which was further used in the analysis plot #3.
3.  The aware_covid dataset was combined with happiness dataset in this part by the column country using an 'inner_join' function.
4.  The final result of the combination was the aware_covid_happiness dataset.

# Exploratory Data Analysis

To achieve our goals, we explored the data by visualization specific information into plots, and recognizing its patterns.

The first aspect that we found interesting is how people's happiness based on their climate awareness by looking at how climate change issues aligns with happiness indicators(e.g. life satisfaction, well-being, and economic factors) across the top 5 countries who has highest GDP per capita in the world. shown in \@ref(fig:insight1).

This plot provides an analysis of the relationship between climate awareness and happiness (measured by life_ladder) across countries with high GDP per capita. By examining these factors, we understand how climate awareness correlates with life satisfaction in economically advanced regions.

The x-axis represent the countries with highest GDP per capita and the y-axis indicates life satisfaction where higher values tells greater life satisfaction. The data is split by awareness_level using facet_wrap(), allowing for comparison across different levels of climate awareness where fill colour intesity varies based on the climate awareness score.

The height of each bar represents the life satisfaction score for a particular country based on their awareness level.The fill color, which represents the awareness score, provides a visual cue about the intensity of climate awareness.The faceted display allows us to compare life satisfaction across different awareness levels independently, offering insights into how various awareness levels relate to happiness. This plot suggest trends that could lead to further investigation. For example, if higher awareness levels correlate with higher life satisfaction, it could imply that awareness positively affects quality of life, possibly through environmental values and proactive behaviors.

```{r insight1, fig.cap="This is a figure caption that you will need to change in order to get good marks in the visualization rubric items."}
aware_happiness <- inner_join(climate_awareness_update2, happiness, by = "country")|>
rename( awareness_level = answer)|>
  arrange(desc(log_gdp_per_capita))|>
  head(30)
aware_happiness
plot_aware_happiness<-
 ggplot(aware_happiness, aes(x= country, y= life_ladder, fill = score ))+
  geom_bar(stat =  "identity")+
  geom_smooth( method = "lm")+
  facet_wrap(~awareness_level)+
   
   labs(title = "Happiness of People based on their climate awareness",
        x = "Countries with highest GDP per capita",
        y = "Ranking based on average happiness/ life satisfaction")+
   theme(axis.text.x = element_text(angle = 90))

plot_aware_happiness
```

This insight is supported by the summary statistics in table \@ref(tab:summary_stats)

```{r summary_stats}
summary_stats <- aware_happiness |>
  summarise(
    mean_life_ladder = mean(life_ladder, na.rm = TRUE),
    median_life_ladder = median(life_ladder, na.rm = TRUE),
    sd_life_ladder = sd(life_ladder, na.rm = TRUE),
    
    mean_gdp = mean(log_gdp_per_capita, na.rm = TRUE),
    median_gdp = median(log_gdp_per_capita, na.rm = TRUE),
    sd_gdp = sd(log_gdp_per_capita, na.rm = TRUE),
    
    mean_score = mean(score, na.rm = TRUE),
    median_score = median(score, na.rm = TRUE),
    sd_score = sd(score, na.rm = TRUE)
  )
summary_stats
summary_table<- summary_stats|>
  kable(
    caption = "Summary Statistics",
    col.names = c(
                  "Mean Life_Ladder", "Median Life_Ladder", "SD Life_Ladder", 
                  "Mean Log GDP", "Median Log GDP", "SD Log GDP",
                  "Mean Score", "Median Score", "SD Score")
  ) 
```

The next insight that we found is shown in \@ref(fig:insight2)that gdp per capita increases covid cases tend to increase as well, European countries tend to generate the most capital while simultaneously being the least corrupt and that covid affected Asia and Europe the most.

1.  GDP per capita focuses on the proportion of economic output per person. GDP acts as a measure of economic prosperity of a country and therefore an increase in population over time. This increase of population leads to a larger spread of covid and hence more people contracting the virus. So, as the x variable of GDP per capita increases, the y axis increases too, signifying a positive correlation coefficient.
2.  We hypothesize that European's strong economic output likely stems from the aparrent lack of corruptions. This is because corrupt societies can have major economic problems (i.e not paying citizens proper wages and overly high taxation rates). The lack of corruption thus motivates citizens to work harder, and in turn an increase in economic output. By this logic, if we plotted each countries corruption level (y) vs GDP per capita (x) on a single graph, we would see tha tthe european countries would be far right on the x axis and lower on the y axis than the other continent.
3.  Covid-19 appeared to affect Asian and European countries the greatest. This is likely due to the high populations of European and Asian countries relative to the other countries. This then increases the overall spread of the disease throughout the population and increases the amount of Covid cases.

```{r insight2, fig.height=4, fig.width=6, fig.cap="This is a figure caption that you will need to change in order to get good marks in the visualization rubric items."}
happiness_covid <- covid_2020 |>
  inner_join(happiness, by = "country") 
happiness_covid
happiness_covid_altered <- happiness_covid |>
  group_by(country) |>
  mutate(
    total_covid_cases = sum(total_cases[year %in% c(2020, 2021, 2022, 2023, 2024)], na.rm = TRUE),
    corruption_base = case_when(
      perceptions_of_corruption < 0.3333 ~ "low corruption",
      perceptions_of_corruption >= 0.3333 & perceptions_of_corruption < 0.6666 ~ "medium corruption",
      perceptions_of_corruption >= 0.6666 ~ "high corruption"
    ) ) |>
  ungroup()
  happiness_covid_altered_filtered <- happiness_covid_altered |>
  filter(
    total_covid_cases >= (quantile(total_covid_cases, 0.25) - 1.5*IQR(total_covid_cases)) &
    total_covid_cases <= (quantile(total_covid_cases, 0.75) + 1.5*IQR(total_covid_cases))
  )
happiness_covid_plot <- happiness_covid_altered_filtered |>
  ggplot(aes(x = log_gdp_per_capita, y = total_covid_cases, color = corruption_base)) +
  geom_point(size = 1.2, alpha = 0.7) + 
  scale_color_manual(values = c("low corruption" = "blue", "medium corruption" = "orange", "high corruption" = "red")) +
  labs(
    x = "Log GDP per Capita",
    y = "Total COVID-19 Cases (Log Scale)",
    color = "Corruption Level",
    title = "COVID-19 Total Cases vs. GDP per Capita by Corruption Level",
    caption = "Each point represents a country, colored by corruption level"
  ) +
  facet_wrap(~continent) +
  theme_minimal() 
happiness_covid_plot
```

This insight is supported by the summary statistics in table \@ref(tab:summary_stats).

```{r}
summary_table <- happiness_covid_altered_filtered |>
  group_by(corruption_base) |>
  summarise(
    mean_total_cases = mean(total_covid_cases, na.rm = TRUE),
    median_total_cases = median(total_covid_cases, na.rm = TRUE),
    sd_total_cases = sd(total_covid_cases, na.rm = TRUE),
    mean_log_gdp = mean(log_gdp_per_capita, na.rm = TRUE),
    median_log_gdp = median(log_gdp_per_capita, na.rm = TRUE),
    sd_log_gdp = sd(log_gdp_per_capita, na.rm = TRUE),
    mean_happiness = mean(happiness_score, na.rm = TRUE),
    median_happiness = median(happiness_score, na.rm = TRUE),
    sd_happiness = sd(happiness_score, na.rm = TRUE)
  )
summary_table |>
  kable(
    caption = "Summary Statistics by Corruption Level",
    col.names = c("Corruption Level", 
                  "Mean Total Cases", "Median Total Cases", "SD Total Cases", 
                  "Mean Log GDP", "Median Log GDP", "SD Log GDP",
                  "Mean Happiness", "Median Happiness", "SD Happiness")
  ) 

```

Finally, \@ref(fig:insight3) shows the relationship between people's positive emotional experiences (positive affect column in happiness data) and the number of covid cases per capita (aware_covid data) across different continents. Plotting these two variables will allow us to see whether there is a clear trend that connects the number of cases of covid with higher or lower levels of positive affect. Because the data is displayed independently for every continent, we can see regional variations and determine whether geographic or cultural factors may have an impact on the connection between covid distribution and emotional well-being.

```{r insight3, fig.height=4, fig.width=6, fig.cap="This is a figure caption that you will need to change in order to get good marks in the visualization rubric items."}
aware_covid_happiness <- inner_join(happiness, aware_covid, by="country")|>
  ggplot() +
    aes(x = positive_affect, y = covid_per_capita, colour = continent) +
    geom_point() +
    geom_smooth(se = FALSE, method = "lm", formula = y ~ x) +
    labs(
        x = "Positive Affect",
        y = "Total covid cases per capita",
        title = "Covid Cases versus Positive Affect",
        colour = "Continent"
    )

aware_covid_happiness
```

This insight is supported by the summary statistics in table \@ref(tab:summary_stats)

```{r}

```

# Conclusion and Future Work

Overall, we found that the analysis of the three datasets on happiness, climate awareness, and COVID-19 provides meaningful insights into the interplay between economic status, climate awareness, and public health impacts across different regions.

In conclusion, higher climate awareness may contribute positively to life satisfaction in wealthier nations, while economic prosperity and low corruption seem to provide a buffer against pandemic challenges. Our research shows that corruption undermines public health systems and democratic institutions, damaging governments' ability to respond effectively to crises like COVID-19 [^6].Regional differences in emotional resilience during COVID-19 highlight the role of culture and geography in shaping public response to crises. This could also be explained in the fact that humans are very dependent on social interactions. [^7]

[^6]: <https://www.transparency.org/en/blog/corruption-and-climate-vulnerability-a-devastating-relationship>

[^7]: <https://cifar.ca/cifarnews/2021/03/17/how-has-covid-19-affected-world-happiness/>

The next steps in this analysis would include further exploraon to see if the positive relationship between climate awareness and happiness also holds for countries with lower GDP per capita. This could help determine if the correlation is universal or influenced by economic factors. As well as, examining factors like healthcare access, government spending on health, and vaccination rates to see how these might impact COVID-19 outcomes in high-GDP countries. Finally, adding data on governance quality could provide a more nuanced view of how economic output and low corruption interplay in different regions.

This analysis has some limitations that affect how broadly we can apply the findings. Since we use country-level data, we may miss important differences within large or diverse countries, and because the analysis is based on correlations, we can’t say for sure that one factor directly causes another (like climate awareness leading to higher happiness). By focusing on data from 2020, we capture only the early impacts of the pandemic, which may not show the full, long-term effects. Cultural differences in reporting happiness and emotional responses could mean our results don’t apply equally to all regions. Additionally, the absence of health and economic factors, and the use of self-reported happiness data, introduce some subjectivity into the results. While the analysis provides useful insights, these limitations mean that conclusions should be interpreted carefully.

# References

1.  <https://data.humdata.org/dataset/dc9f2ca4-8b62-4747-89b1-db426ce617a0/resource/6041db5f-8190-47ff-a10b-9841325de841/download/climate_change_opinion_survey_2022_aggregated.xlsx>
2.  <https://dataforgood.facebook.com/dfg/tools/climate-change-opinion-survey>
3.  <https://covid.ourworldindata.org/data/owid-covid-data.csv>
4.  <https://worldhappiness.report/>
5.  <https://worldhappiness.report/ed/2023/#appendices-and-data>
6.  <https://www.transparency.org/en/blog/corruption-and-climate-vulnerability-a-devastating-relationship>
7.  <https://cifar.ca/cifarnews/2021/03/17/how-has-covid-19-affected-world-happiness/>
